---
title: "Texas Census Block Groups by Racial Majority"
author: "Kaitlan Wong"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidycensus)
library(tidyverse)
library(readxl)
library(writexl)
library(dplyr)
```

```{r}
# Look at ACS 5-Year tables

v21 <- load_variables(2021, "acs5", cache = TRUE)

View(v21)
```

***Load 2021 ACS 5-Year Estimates using the Hispanic or Latino by Race tables.***

Variable definitions: (from "HISPANIC OR LATINO ORIGIN BY RACE" concept in v21)

-   **B03002_001**: Estimate!!Total:

-   **B03002_003**: Estimate!!Total:!!Not Hispanic or Latino:!!White alone

-   **B03002_004**: Estimate!!Total:!!Not Hispanic or Latino:!!Black or African American alone

-   **B03002_005**: Estimate!!Total:!!Not Hispanic or Latino:!!American Indian and Alaska Native alone

-   **B03002_006**: Estimate!!Total:!!Not Hispanic or Latino:!!Asian alone

-   **B03002_007**: Estimate!!Total:!!Not Hispanic or Latino:!!Native Hawaiian and Other Pacific Islander alone

-   **B03002_008**: Estimate!!Total:!!Not Hispanic or Latino:!!Some other race alone

-   **B03002_009**: Estimate!!Total:!!Not Hispanic or Latino:!!Two or more races:

-   **B03002_012**: Estimate!!Total:!!Hispanic or Latino:

See <https://walker-data.com/tidycensus/articles/basic-usage.html>

```{r}
# Get ACS data
acs21 <- get_acs(
  state = "TX",
  geography = "block group",
  variables = c("B03002_001", "B03002_003", "B03002_004", "B03002_005", "B03002_006","B03002_007", "B03002_008", "B03002_009", "B03002_012"),
  survey = "acs5",
  year = 2021,
  recode = TRUE,
)

# View data
head(acs21)
```

Reshape the data.

```{r}
# Remove the "moe" column from the dataset
acs21_no_moe <- acs21 %>%
  select(-moe)

library(tidyr)

# Reshape the data and aggregate
data21 <- acs21_no_moe %>%
  pivot_wider(names_from = variable, values_from = estimate) %>%
  group_by(NAME) %>%
  ungroup()

# View data
head(data21)

```

```{r}
# Create AAPI and Other columns and rename variables

names <- data21 %>%
  mutate(
    NH_AAPI = B03002_006 + B03002_007,
    NH_Other = B03002_008 + B03002_009
  ) %>%
  rename(
    Total_Pop = B03002_001,
    NH_White = B03002_003,
    NH_Black = B03002_004,
    NH_AIAN = B03002_005,
    NH_Asian = B03002_006,
    NH_NHPI = B03002_007,
    NH_SomeOtherAlone = B03002_008,
    NH_Multiracial = B03002_009,
    Hispanic = B03002_012
  )

head(names)
```

```{r}
# Calculate percentage columns
percentage_data <- names %>%
  mutate(
    NH_White_Per = NH_White / Total_Pop,
    NH_Black_Per = NH_Black / Total_Pop,
    NH_AIAN_Per = NH_AIAN / Total_Pop,
    NH_Asian_Per = NH_Asian / Total_Pop,
    NH_NHPI_Per = NH_NHPI / Total_Pop,
    NH_SomeOtherAlone_Per = NH_SomeOtherAlone / Total_Pop,
    NH_Multiracial_Per = NH_Multiracial / Total_Pop,
    Hispanic_Per = Hispanic / Total_Pop,
    NH_AAPI_Per = NH_AAPI / Total_Pop,
    NH_Other_Per = NH_Other / Total_Pop
  )

# View the data with percentage columns
head(percentage_data)
```

```{r}
# Calculate the racial majority column
racial_majorities <- percentage_data %>%
  mutate(
    Racial_Majority = case_when(
      NH_White_Per > 0.5 ~ "NH_White",
      NH_Black_Per > 0.5 ~ "NH_Black",
      NH_AIAN_Per > 0.5 ~ "NH_AIAN",
      NH_AAPI_Per > 0.5 ~ "NH_AAPI",
      NH_Other_Per > 0.5 ~ "NH_Other",
      Hispanic_Per > 0.5 ~ "Hispanic",
      TRUE ~ "Racially Diverse"
    )
  )

# View the data with the racial majority column
head(racial_majorities)
```

Save data to Excel.

```{r}
# Define the Excel file name
excel_file_name <- "Texas Census Block Groups by Racial Majority_2021 ACS5.xlsx"

# Export the entire data frame to Excel
write_xlsx(racial_majorities, excel_file_name)
```
